---
title: "Simulations"
output: github_document
---

```{r setup}
library(tidyverse)
library(rvest)

set.seed(1)
```

## Mean and SD for one n

```{r}
sim_mean_sd = function(n, mu = 2, sigma = 3) {
  
  sim_data = tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )
  
  sim_data |> 
    summarize(
      mu_hat = mean(x),
      sigma_hat = sd(x)
    )
}
```

let's run it a 100 times "simulate" with a `for` loop
```{r}
output = vector("list", 100)

for (i in 1:100) {
  output[[i]] = sim_mean_sd(30)
}

sim_results = bind_rows(output)
```

```{r}
sim_results_df = 
  expand_grid(
    sample_size = 30,
    iter = 1:100
  ) |> 
  mutate(
    estimate_df = map(sample_size, sim_mean_sd)
  ) |> 
  unnest(estimate_df)
```


 
```{r}
sim_results_df |> 
  ggplot(aes(x = mu_hat)) + 
  geom_density()
```
 
```{r}
sim_results_df |> 
  pivot_longer(
    mu_hat:sigma_hat,
    names_to = "parameter", 
    values_to = "estimate") |> 
  group_by(parameter) |> 
  summarize(
    emp_mean = mean(estimate),
    emp_sd = sd(estimate)) |> 
  knitr::kable(digits = 3)
```

## Several N's

```{r}
sim_results_df =   
  map(1:100, \(i) sim_mean_sd(30, 2, 3)) |> 
  bind_rows()

```

```{r}
sim_results_df = 
  expand_grid(
    sample_size = c(30, 60, 120, 240),
    iter = 1:1000
  ) |> 
  mutate(
    estimate_df = map(sample_size, sim_mean_sd)
  ) |> 
  unnest(estimate_df)
```
Instead of fixing the sample size, you’re varying it systematically.

You can now compare how estimator variability changes as sample size increases.

This shows an important principle: larger n → smaller variance of the estimator.

## Plots

```{r}
sim_results_df |> 
  mutate(
    sample_size = str_c("n = ", sample_size),
    sample_size = fct_inorder(sample_size)) |> 
  ggplot(aes(x = sample_size, y = mu_hat, fill = sample_size)) + 
  geom_violin()
```

Empiral mean and variance of estimates 
```{r}
sim_results_df |> 
  pivot_longer(
    mu_hat:sigma_hat,
    names_to = "parameter", 
    values_to = "estimate") |> 
  group_by(parameter, sample_size) |> 
  summarize(
    emp_mean = mean(estimate),
    emp_var = var(estimate)) |> 
  knitr::kable(digits = 3)
```

## Regression

```{r}
sim_regression = function(n, beta0 = 2, beta1 = 3) {
  
  sim_data = 
    tibble(
      x = rnorm(n, mean = 1, sd = 1),
      y = beta0 + beta1 * x + rnorm(n, 0, 1)
    )
  
  ls_fit = lm(y ~ x, data = sim_data)
  
  tibble(
    beta0_hat = coef(ls_fit)[1],
    beta1_hat = coef(ls_fit)[2]
  )
}
```

```{r}
sim_results_df = 
  expand_grid(
    sample_size = 30,
    iter = 1:500
  ) |> 
  mutate(
    estimate_df = map(sample_size, sim_regression)
  ) |> 
  unnest(estimate_df)
```

```{r}
sim_results_df |> 
  ggplot(aes(x = beta0_hat, y = beta1_hat)) + 
  geom_point()
```

## two parameters

sd and sample size

```{r}
sim_results_df = 
  expand_grid(
    sample_size = c(30, 60, 120, 240),
    true_sd = c(6, 3),
    iter = 1:1000
  ) |> 
  mutate(
    estimate_df = 
      map2(sample_size, true_sd, \(n, sd) sim_mean_sd(n = n, sigma = sd))
  ) |> 
  unnest(estimate_df)
```

```{r}
sim_results_df |> 
  mutate(
    true_sd = str_c("True SD: ", true_sd),
    true_sd = fct_inorder(true_sd),
    sample_size = str_c("n = ", sample_size),
    sample_size = fct_inorder(sample_size)) |> 
  ggplot(aes(x = sample_size, y = mu_hat, fill = sample_size)) + 
  geom_violin() + 
  facet_grid(. ~ true_sd)
```

